### 2013-03-01 branching-transit.tex
WorkingWiki importation

\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\var}{\mathop{\mbox{Var}}}
\newcommand{\Exp}{\mathop{\mbox{Exp}}}

@@

>>>>

\newcommand{\E}{\mathbb{E}}
\renewcommand{\P}{\mathbb{P}}
\newcommand{\var}{\mathop{\mbox{Var}}}
\newcommand{\Exp}{\mathop{\mbox{Exp}}}
\newcommand{\migrate}{\lambda_\text{mig}}

@@

\subsection{Occupation density}

Suppose we have branching processes beginning at the origin at some constant rate $\lambda$ in time.
Then the occupation density at distance $x$, i.e.\ the expected number of particles found at location $x$ at any given time, at stationarity,
is $\lambda \int_0^\infty \E[Z_t(dx)] dt$, where $Z_t$ is the branching process.
Since spatial motion is independent of branching,
\begin{align}
  \E[Z_t(dx)] = e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) ,
\end{align}
and so using G\&R 3.471.9 and 8.468 again,
\begin{align} \label{eqn:bp_occupation}
  \int_0^\infty \E[Z_t(dx)] dt  &= \int_0^\infty e^{-\alpha t} \frac{1}{(2\pi \sigma^2 t)^{d/2}} \exp\left(-\frac{x^2}{2\sigma^2 t} \right) dt \\
  &= \begin{cases}
    \frac{ 2^{1/4} x^{1/2} }{ \sqrt{\pi} \sigma^{3/2} \alpha^{1/4} } K_{1/2} \left(\frac{x \sqrt{2\alpha}}{\sigma}\right) =
    \frac{1}{\sigma \sqrt{2 \alpha}} \exp\left( - \frac{ \sqrt{2\alpha} x }{ \sigma } \right) \qquad & d=1 \\
    \frac{1}{\sigma^2 \pi} K_0\left(\frac{x \sqrt{2\alpha}}{\sigma}\right) \sim \frac{ e^{-\frac{x\sqrt{2\alpha}{\sigma}}} }{ \sqrt{ \sigma^3  2\pi x \qquad & d=2  ,
\end{cases} 
\end{align}
where $K_0(z)$ is the modified Bessel function of the second kind of order 0.
Note that this is the Green function for a Brownian motion killed at rate $\alpha$.

A more back-of-the-envelope calculation of this would be:
we know that the probability a branching process arrives at the patch is given by $(1/\mu_K)\exp(-x\sqrt{2 \alpha}/\sigma)$;
if it does it is present in $\mu_K$ copies, each of which leave a total of $1/s_m$ offspring.
To translate this to the above we need the typical area covered by these (in some sense?), 
which must be around $\sigma \sqrt{2 \alpha}$.

\subsection{An integral}

For reference, by G\&R 3.471.9 and 8.468,
\begin{align}
  \int_0^\infty x^{n+1/2} e^{-\frac{\beta}{x}-\gamma x} dx &= 2 \left( \frac{\beta}{\gamma} \right)^{\frac{2n+1}{4}} K_{n+3/2}(2\sqrt{\beta\gamma}) \quad \mbox{(for general $n$)} \\
  \mbox{($n$ an integer)} \quad &= \sqrt{\pi} \beta^{\frac{n}{2}} \gamma^{-\frac{n+1}{2}} e^{-2\sqrt{\beta\gamma}} \sum_{k=1}^n \frac{ (n+k)! }{ k! (n-k)! (4\sqrt{\beta \gamma})^k } .
\end{align}

Note that (8.451.6) $K_\nu(z) = \sqrt{\frac{\pi}{2z}} e^{-z} (1 + O(1/z))$.

\section{The big picture}

Here are notes to clarify what's going on for writing the main paper.

We can calculate the rate of ``core'' establishment, i.e.\ the mean number of reproductive events that lead to establishment as defined below, in two ways:
\begin{align}
  (\mbox{rate of establishment}) &= (\mbox{outflux of BPs from old patch}) \times (\mbox{\# estab events per BP}) \\
    \begin{split} &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
  & \qquad \qquad \times (\mbox{\# estab events per successful BP}) \end{split} \\
  \begin{split} &= (\mbox{subcritical occupation density}) \times \rho \times (\mbox{\# offspring per subcritical indiv}) \\
    & \qquad \qquad \times (\mbox{prob of estab in the patch}) . \end{split} 
\end{align}
The probability of establishment of a branching process is in equation~\eqref{eqn:bp_prob_estab};
the subcritical occupation density is the integral of equation~\eqref{eqn:eqfreq} over the patch;
the number of (extra) offspring per subcritical individual is $s_B+s_m$;
and the probability of establishment in the patch is $2s_B/\xi^2$.
The ``outflux of branching processes'' we have not computed;
but we have compute the occupation density from this branching process approximation above,
so equating ($\rho$ times $C\exp(-\sqrt{2 s_m}x/\sigma)$) and (outflux times equation~\eqref{eqn:bp_occupation})
we have that the effective outflux (from the edge of the already occupied patch!) is 
\begin{align}
  (\mbox{outflux of BPs from old patch}) = \frac{ C \rho \sqrt{\alpha} }{ \sqrt{\pi} } ,
\end{align}
rather than $C\rho$ which we might have guessed.
The number of establishment events per BP is as follows, in averages: 
with prob $1/\mu_K \exp(-\sqrt{\alpha} x/ \sigma)$ a total of $\mu_K$ indivs arrive at the new patch; 
each produce $1/s_m$ total core offspring;
each of these produce $s_B+s_m$ extra offspring which each have chance $2s_B/\xi^2$ of establishing,
i.e.
\begin{align}
  (\mbox{\# estab events per BP}) = \frac{(s_B+s_m) 2 s_B }{ s_m \xi^2 } \exp\left(-\frac{\sqrt{\alpha} x}{\sigma}\right) .
\end{align}
\textbf{Now} note that what we \textbf{actually} want to compute is not this rate, but rather 
\begin{align}
  \migrate &= (\mbox{outflux of BPs from old patch}) \times (\mbox{prob of estab of a distant BP}) \\
  &= C \rho \sqrt{\alpha} \times \frac{2 s_b}{\sqrt{\pi} \xi^2} \exp\left(- \frac{\sqrt{\alpha}x}{\sigma} \right) .
\end{align}
Finally, combining these, we can also compute that
\begin{align}
  (\mbox{\# estab events per successful BP}) = \frac{s_B+s_m}{ \sqrt{s_m} } .
\end{align}

### 2013-03-01 patchy-selection-paper.tex
WorkingWiki importation

Due to migration--selection balance, 
there will be a small number of $B$ alleles descended from those of the first patch
present in the regions where they are disadvantageous;
the expected frequency at distance $r$ may again be expressed in terms of Jacobi elliptic functions \citep[Appendix XXX][]{barton},
although the equations are again not particularly helpful.
This expected frequency is the \emph{time-averaged} occupation frequency,
i.e.\ the average frequency of the allele at each location over a long time period,
below we study the fluctuations about this mean.
More helpful than the exact expressions,
we know that if $q(r)$ is the expected proportion present at migration--selection balance at distance $r$ from the first patch
in the absence of the second patch, 
then for large $r$ this proportion decays exponentially, 
\begin{align} \label{eqn:eqfreq}
  q(r) \approx C \exp( -\sqrt{s_m} r / \sigma),
\end{align}
where $C$ is a constant depending on the geographic shape of the populations and the selection coefficients. 
(Probably $C=1/2$ is close enough for applications; see Appendix XXX.)
\plr{same appendix with sims as above}
\gc{Should we also use this approx. to get the approx. rate of establishing mutations
arising outside of the patch, above. Feel like in the mutation rate
part we could pull $s_B$ out front of the rate as well.} 
\plr{huh?}

This allows us to get a good sense of the time scale of adaptation by migration between separated patches, as follows.
First imagine that the new allele is not advantageous in the second patch,
so that the expected number of alleles present in a small area $\Delta$ at distance $r$ from the already established patch
is $\Delta \rho q(r)$.
As these copies are disadvantageous, they form a subcritical branching process, which die out fairly soon.
Now add to this local selective advantage by stipulating that any copy of the allele in the new patch
has ``extra'' offspring that are marked in some way, on average $(1+s_B)-(1-s_m)=s_B+s_m$ additional offspring per generation.
If the probability of establishment \emph{in the new patch} is $p(x)$ as above,
then each of these new, marked copies appearing at location $x$ has probability $p(x)$ of establishing.
(This general relation is stated formally in Appendix XXX.)
Therefore, the total rate of establishment 
(i.e.\ the mean number of successfully establishing migrant lineages per generation)
is $\rho (s_B+s_m) \int q(|x|) p(x) dx$.
Using a similar approximation as in section \ref{ss:patchymutation},
we can approximate this, for a patch of area $A$ at distance $R$ from an already occupied patch, as
\begin{align}
  \migrate(R) = \rho A \frac{s_B}{\xi^2} (s_B+s_m) \exp\left( -\frac{ \sqrt{s_m} R}{\sigma} \right),
\end{align}
where we have set $C=1/2$ in equation \eqref{eqn:eqfreq}.
Using different methods, we arrive at the same functional form below in section \ref{ss:haplotype_length}.
(Note that if the size of the new patch is large relative to $\sqrt{s_m}/\sigma$,
then the integral of $q(x)$ over the patch will not be well approximated by $A \times q(R)$,
but the adjustment is easy to make.)

@@

Taking expressions \eqref{eqn:mutrate} and \eqref{eqn:migrate} above,
we see that if the patches each have sufficiently large area,
the rate of migration and mutation will be on the same order if 
$\mu \approx ((s_B+s_m)/2) \exp(-\sqrt{s_m} R / \sigma)$,
where $R$ is the distance between the patches.
(Note that in both cases, $A$ is the area of the not-yet-adapted patch.)

@@

If we are willing to take these approximations at face value, 
and assuming both rates are small, 
the time till the first patch is colonized by the $B$ allele will be approximately exponentially distributed with rate $2 \mutrate$.
Following this, the time till the second patch is subsequently colonized 
(via either migration or new mutation) 
will be exponentially distributed with rate $\mutrate+\migrate$.
Both these times scale linearly with $\xi^2/(\rho A s_{B})$, so that 
the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
and decrease with patch population size and benefical seleciton coefficient.
Furthermore, the probability that the second adaptation is a new mutation,
i.e.\ the probability of parallel adapatation, is
\begin{equation}
  \frac{\mutrate}{\mutrate+\migrate} = \frac{2\mu/(s_B+s_m)}{2\mu/(s_B+s_m) + \exp\left(- \sqrt{s_m} R / \sigma \right) },
\end{equation}
so that the probability of parallel mutation should increase
approximately logistically with the distance between the patches at rate $\sqrt{s_m} /\sigma$.

@@

If not yet adapted,
a patch of area $A$ will acquire the adaptation by new mutation at rate $2 s_B \mu \rho A/\xi^2$.
Without loss of generality, assume that patches $1, \ldots, k$ are already adapted,
and these patches are at distances $R_1, \ldots, R_k$ away from this unadapted patch, 
then the total rate of adaptation through migration is
\begin{equation}
  \frac{ s_B (s_B+s_m) \rho A}{\xi^2} \sum_{i=1}^{k} \exp\left(- \sqrt{s_m} R_i/\sigma\right),
\end{equation}
and the probability that the adaptation is from patch $i$ is $\exp\left(- \sqrt{s_m} R_i/\sigma\right)/\sum_{j=1}^{k} \exp\left(- \sqrt{s_m} R_j/\sigma\right)$.
These specify the transition rates of the process.

@@

Now we can add in the spatial motion, by following only the motion of the ``trunk''
(i.e.\ the red line in figure \ref{fig:branching_decomp}).
We will first work in one dimension.
If the variance of the parent-offspring distance is $\sigma^2$ (and the distribution is not too heavy-tailed),
then the motion of this is approximately Brownian, with variance $\sigma^2$ per generation.
Suppose that the new patch is at distance $x$.
The probability density for the first time such a Brownian motion $B_s$ hits $x$ is
\citep[XXX]{feller}
% note $\sigma B_t \deq B_{\sigma^2 t}$ so $\P\{ \sup_{s\le t} \sigma B_s \le x \} = \P\{ \sup_{s \le \sigma^2 t} B_s \le x \}
\begin{align} \label{eqn:hitting_dist}
  \P\{ B_t=x \;\mbox{and} \; B_u<x \;\mbox{for all}\; u<t\} =  \frac{x}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t\sigma^2}\right) .
\end{align}
The chance that the descendants of a given migrant arrives at the new patch at time $t$
is now just equation~\eqref{eqn:hitting_dist} multiplied by the chance that 
the migrant family is still alive at time $t$.
For convenience, we define $\alpha = - \log(1-s_m) \approx s_m$,
so that $1-p_e(t) \simeq (1/\mu_K) \exp(-\alpha t)$,
and define $\mu_K$ to be the limiting mean family size conditioned on survival.
By integrating over $t$, we get the chance that descendants of the migrant ever arrive at the new patch is
\begin{align} \label{eqn:estab_prob}
  & \int_0^\infty \frac{(1-s_m)^t}{\mu_K} \frac{x}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t\sigma^2}\right)  dt \\
  & \qquad = \frac{1}{\mu_K} \exp\left( - \frac{ x \sqrt{\alpha}}{\sigma} \right) .
\end{align}
(Note that this is, up to the constant $\mu_K$, the chance that a Brownian motion killed at rate $\alpha$
ever hits $x$ \citep{feller}.
This approximation holds if $x$ is large enough that the approximation $1-p_e(t) \approx (1-s_m)^t/\mu_K$ holds.)

@@

The probability of establishment of a single migrant in a patch distance $x$ away
is equal to the probability that the migrant's family makes it to the patch at all,
multiplied but the chance that it establishes locally, given that it makes it there.
We have computed the first term in equation~\eqref{eqn:estab_prob},
and we make the small assumption that the second term is just the probability of local fixation
of a beneficial allele present in $K$ copies,
i.e.~$\E[1-(1-p_f(s_b))^K] \approx 2 s_b \mu_K/\xi^2$,
where $p_f(s_b) \approx 2s_b/\xi^2$ is the probability of fixation of a single mutant with additive benefit $s_b$.
Therefore, the probability of establishment is
\begin{align}
  F(x) &\approx \frac{2s_b}{\xi^2} \exp\left( - \frac{\sqrt{\alpha} x}{\sigma} \right) ,
\end{align}
which verifies the less intuitive calculation of $\migrate$ above.

@@

In the text at several points appear integrals of the form
\begin{align}
  \int_0^\infty t^c \exp \left( - \alpha t^d - \beta t^{d+1} \right) dt 
\end{align}
where $c$ is a positive integer, $d$ is the dimension, and $\alpha$ and $\beta$ are positive real numbers.
This could be evaluated through standard numerical methods; below we describe a power series expansion.
Changing variables to $u = \beta t^{d+1}$, this becomes
\begin{align}
    \left( (d+1)^{-1} \beta^{ (1-c)/(d+1) } \right) \int_0^\infty u^{(c+1-d)/(d+1)} \exp\left( - \alpha \beta^{-d/(d+1)} u^{d/(d+1)} - u \right) du ,
\end{align}
so it suffices to evaluate the function
\begin{equation}
    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt ,
\end{equation}
in the case that $a=(c-d+1)/(d+1)$, $b=d/(d+1)$, and $x$ is a function of the demographic paramters.
Since $a$ and $b$ only depend on the dimension and the quantity being computed,
we are interested in $G$ as a function of $x$.
At least in the case $b=d/(d+1)$ it is possible to express $G$ as a finite sum of gamma functions,
but we proceed with a simpler method.
Note that $\partial_x G(a,b,x) = -G(a+b,b,x)$,
and that at $x=0$, the function $G$ is the gamma function $G(a,b,0) = \Gamma(a+1)$.
Therefore, a Taylor series for $G$ would be
\[
    G(a,b,x) = \sum_{n \ge 0} = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
\]
It is easy to check using Stirling's formula that $\limsup_{n \to \infty} ( x^n \Gamma(a+nb+1)/n! )^{1/n} = 0$
if $b<1$, so the sum converges.

>>>>

Due to migration--selection balance, 
there will be a small number of $B$ alleles descended from those of the first patch
present in the regions where they are disadvantageous;
the expected frequency at distance $r$ may again be expressed in terms of Jacobi elliptic functions \citep[Appendix XXX][]{barton},
although the equations are again not particularly helpful.
This expected frequency is the \emph{time-averaged} occupation frequency,
i.e.\ the average frequency of the allele at each location over a long time period,
below we study the fluctuations about this mean.
More helpful than the exact expressions,
we know that if $q(r)$ is the expected proportion present at migration--selection balance at distance $r$ from the first patch
in the absence of the second patch, 
then for large $r$ this proportion decays exponentially, 
\begin{align} \label{eqn:eqfreq}
  q(r) \approx C \exp( -\sqrt{2 s_m} r / \sigma),
\end{align}
where $C$ is a constant depending on the geographic shape of the populations and the selection coefficients. 
(Probably $C=1/2$ is close enough for applications; see Appendix XXX.)
\plr{same appendix with sims as above}
\gc{Should we also use this approx. to get the approx. rate of establishing mutations
arising outside of the patch, above. Feel like in the mutation rate
part we could pull $s_B$ out front of the rate as well.} 
\plr{huh?}

This allows us to get a good sense of the time scale of adaptation by migration between separated patches, as follows.
First imagine that the new allele is not advantageous in the second patch,
so that the expected number of alleles present in a small area $\Delta$ at distance $r$ from the already established patch
is $\Delta \rho q(r)$.
As these copies are disadvantageous, they form a subcritical branching process, which die out fairly soon.
Now add to this local selective advantage by stipulating that any copy of the allele in the new patch
has ``extra'' offspring that are marked in some way, on average $(1+s_B)-(1-s_m)=s_B+s_m$ additional offspring per generation.
If the probability of establishment \emph{in the new patch} is $p(x)$ as above,
then each of these new, marked copies appearing at location $x$ has probability $p(x)$ of establishing.
(This general relation is stated formally in Appendix XXX.)
Therefore, the total rate of establishment 
(i.e.\ the mean number of successfully establishing migrant lineages per generation)
is $\rho (s_B+s_m) \int q(|x|) p(x) dx$.
Using a similar approximation as in section \ref{ss:patchymutation},
we can approximate this, for a patch of area $A$ at distance $R$ from an already occupied patch, as
\begin{align}
  \migrate(R) = \rho A \frac{s_B}{\xi^2} (s_B+s_m) \exp\left( -\frac{ \sqrt{2 s_m} R}{\sigma} \right),
\end{align}
where we have set $C=1/2$ in equation \eqref{eqn:eqfreq}.
Using different methods, we arrive at the same functional form below in section \ref{ss:haplotype_length}.
(Note that if the size of the new patch is large relative to $\sqrt{s_m}/\sigma$,
then the integral of $q(x)$ over the patch will not be well approximated by $A \times q(R)$,
but the adjustment is easy to make.)

@@

Taking expressions \eqref{eqn:mutrate} and \eqref{eqn:migrate} above,
we see that if the patches each have sufficiently large area,
the rate of migration and mutation will be on the same order if 
$\mu \approx ((s_B+s_m)/2) \exp(-\sqrt{2 s_m} R / \sigma)$,
where $R$ is the distance between the patches.
(Note that in both cases, $A$ is the area of the not-yet-adapted patch.)

@@

If we are willing to take these approximations at face value, 
and assuming both rates are small, 
the time till the first patch is colonized by the $B$ allele will be approximately exponentially distributed with rate $2 \mutrate$.
Following this, the time till the second patch is subsequently colonized 
(via either migration or new mutation) 
will be exponentially distributed with rate $\mutrate+\migrate$.
Both these times scale linearly with $\xi^2/(\rho A s_{B})$, so that 
the overall time to adaptation is robustly predicted to increase with offspring variance ($\xi^2$)
and decrease with patch population size and benefical seleciton coefficient.
Furthermore, the probability that the second adaptation is a new mutation,
i.e.\ the probability of parallel adapatation, is
\begin{equation}
  \frac{\mutrate}{\mutrate+\migrate} = \frac{2\mu/(s_B+s_m)}{2\mu/(s_B+s_m) + \exp\left(- \sqrt{2 s_m} R / \sigma \right) },
\end{equation}
so that the probability of parallel mutation should increase
approximately logistically with the distance between the patches at rate $\sqrt{s_m} /\sigma$.

@@

If not yet adapted,
a patch of area $A$ will acquire the adaptation by new mutation at rate $2 s_B \mu \rho A/\xi^2$.
Without loss of generality, assume that patches $1, \ldots, k$ are already adapted,
and these patches are at distances $R_1, \ldots, R_k$ away from this unadapted patch, 
then the total rate of adaptation through migration is
\begin{equation}
  \frac{ s_B (s_B+s_m) \rho A}{\xi^2} \sum_{i=1}^{k} \exp\left(- \sqrt{2 s_m} R_i/\sigma\right),
\end{equation}
and the probability that the adaptation is from patch $i$ is $\exp\left(- \sqrt{2 s_m} R_i/\sigma\right)/\sum_{j=1}^{k} \exp\left(- \sqrt{2 s_m} R_j/\sigma\right)$.
These specify the transition rates of the process.

@@

Now we can add in the spatial motion, by following only the motion of the ``trunk''
(i.e.\ the red line in figure \ref{fig:branching_decomp}).
We will first work in one dimension.
If the variance of the parent-offspring distance is $\sigma^2$ (and the distribution is not too heavy-tailed),
then the motion of this is approximately Brownian, with variance $\sigma^2$ per generation.
Suppose that the new patch is at distance $x$.
The probability density for the first time such a Brownian motion $B_s$ hits $x$ is
\citep[XXX]{feller}
% note $\sigma B_t \deq B_{\sigma^2 t}$ so $\P\{ \sup_{s\le t} \sigma B_s \le x \} = \P\{ \sup_{s \le \sigma^2 t} B_s \le x \}
\begin{align} \label{eqn:hitting_dist}
  \P\{ B_t=x \;\mbox{and} \; B_u<x \;\mbox{for all}\; u<t\} =  \frac{x}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t\sigma^2}\right) .
\end{align}
The chance that the descendants of a given migrant arrives at the new patch at time $t$
is now just equation~\eqref{eqn:hitting_dist} multiplied by the chance that 
the migrant family is still alive at time $t$.
For convenience, we define $\alpha = - \log(1-s_m) \approx s_m$,
so that $1-p_e(t) \simeq (1/\mu_K) \exp(-\alpha t)$,
and define $\mu_K$ to be the limiting mean family size conditioned on survival.
By integrating over $t$, we get the chance that descendants of the migrant ever arrive at the new patch is
\begin{align} 
  & \int_0^\infty \frac{(1-s_m)^t}{\mu_K} \frac{x}{\sigma^3 t^{3/2}\sqrt{2\pi}} \exp\left(-\frac{x^2}{2t\sigma^2}\right)  dt \\
  & \qquad = \frac{1}{\mu_K} \exp\left( - \frac{ x \sqrt{2 \alpha}}{\sigma} \right) .\label{eqn:estab_prob}
\end{align}
(Note that this is, up to the constant $\mu_K$, the chance that a Brownian motion killed at rate $\alpha$
ever hits $x$ \citep{feller}.
This approximation holds if $x$ is large enough that the approximation $1-p_e(t) \approx (1-s_m)^t/\mu_K$ holds.)

@@

The probability of establishment of a single migrant in a patch distance $x$ away
is equal to the probability that the migrant's family makes it to the patch at all,
multiplied but the chance that it establishes locally, given that it makes it there.
We have computed the first term in equation~\eqref{eqn:estab_prob},
and we make the small assumption that the second term is just the probability of local fixation
of a beneficial allele present in $K$ copies,
i.e.~$\E[1-(1-p_f(s_b))^K] \approx 2 s_b \mu_K/\xi^2$,
where $p_f(s_b) \approx 2s_b/\xi^2$ is the probability of fixation of a single mutant with additive benefit $s_b$.
Therefore, the probability of establishment is
\begin{align} \label{eqn:bp_prob_estab}
  F(x) &\approx \frac{2s_b}{\xi^2} \exp\left( - \frac{\sqrt{2 \alpha} x}{\sigma} \right) ,
\end{align}
which verifies the less intuitive calculation of $\migrate$ above.

@@

In the text at several points appear integrals of the form
\begin{align}
  \int_0^\infty t^c \exp \left( - \alpha t^d - \beta t^{d+1} \right) dt 
\end{align}
where $c$ is a positive integer, $d$ is the dimension, and $\alpha$ and $\beta$ are positive real numbers.
This could be evaluated through standard numerical methods; below we describe a power series expansion.
Changing variables to $u = \beta t^{d+1}$, this becomes
\begin{align}
    \left( (d+1)^{-1} \beta^{ (1-c)/(d+1) } \right) \int_0^\infty u^{(c+1-d)/(d+1)} \exp\left( - \alpha \beta^{-d/(d+1)} u^{d/(d+1)} - u \right) du ,
\end{align}
so it suffices to evaluate the function
\begin{equation}
    G(a,b,x) := \int_0^\infty  t^a \exp\left( -x t^b - t \right) dt ,
\end{equation}
in the case that $a=(c-d+1)/(d+1)$, $b=d/(d+1)$, and $x$ is a function of the demographic paramters.
Since $a$ and $b$ only depend on the dimension and the quantity being computed,
we are interested in $G$ as a function of $x$.
At least in the case $b=d/(d+1)$ it is possible to express $G$ as a finite sum of gamma functions,
but we proceed with a simpler method.
Note that $\partial_x G(a,b,x) = -G(a+b,b,x)$,
and that at $x=0$, the function $G$ is the gamma function $G(a,b,0) = \Gamma(a+1)$.
Therefore, a Taylor series for $G$ would be
\[
    G(a,b,x) = \sum_{n \ge 0} \frac{(-x)^n}{n!} \Gamma(a+nb+1) .
\]
It is easy to check using Stirling's formula that $\limsup_{n \to \infty} ( x^n \Gamma(a+nb+1)/n! )^{1/n} = 0$
if $b<1$, so the sum converges.
